---
title: "Practical Machine Learning - Project Submission"
author: "mpekari"
date: "Friday, May 22, 2015"
output: html_document
---

This document is the project submission for Coursera.org Practical Machine Learning course. 

I used the caret package and random forest method for fitting the model. 

```{r eval=FALSE}
library(caret)
library(randomForest)
```

I loaded the datasets and marked missing values as NA. 

```{r eval=FALSE}
trainingSet <- read.csv(file = "pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testSet <- read.csv(file = "pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

I removed the first 7 columns (username, timestamps, etc.) as they did not have sensor data. I also removed columns that had mostly NA values. 

```{r eval=FALSE}
trainingSet <- trainingSet[,-(1:7)]
na_columns <- apply(is.na(trainingSet), 2, sum) > 19000
trainingSet <- trainingSet[,!na_columns]
```

I selected only 50% of the training data for fitting the model, so that the fitting would complete in reasonable time with my computer. 

```{r eval=FALSE}
set.seed(4378)
partition <- createDataPartition(y=trainingSet$classe,p=0.5,list=FALSE)
trainingSet.small <- trainingSet[partition,]
```

I fitted the model using the random forest method and 5-fold cross-validation. I used only 5 folds for performance reasons.

```{r eval=FALSE}
modelFit <- train(classe~.,data=trainingSet.small,method="rf",
                trControl=trainControl(method="cv",number=5, verboseIter=TRUE),
                prox=TRUE)
```

The final model has about 0.9865 accuracy.

```{r eval=FALSE}
print(modelFit)
```

```{r echo=FALSE}
cat("Random Forest 

9812 samples
  52 predictor
   5 classes: 'A', 'B', 'C', 'D', 'E' 

No pre-processing
Resampling: Cross-Validated (5 fold) 

Summary of sample sizes: 7851, 7850, 7849, 7849, 7849 

Resampling results across tuning parameters:

  mtry  Accuracy   Kappa      Accuracy SD  Kappa SD   
   2    0.9850191  0.9810448  0.004420207  0.005595374
  27    0.9865475  0.9829826  0.003575824  0.004521962
  52    0.9799229  0.9746009  0.003970556  0.005024378

Accuracy was used to select the optimal model using  the largest value.
The final value used for the model was mtry = 27.
")
```

The out of sample error rate estimated using the the 5 folds is 1.05%.   

```{r eval=FALSE}
print(modelFit$finalModel)
```

```{r echo=FALSE}

cat("Call:
 randomForest(x = x, y = y, mtry = param$mtry, proximity = TRUE) 
               Type of random forest: classification
                     Number of trees: 500
No. of variables tried at each split: 27

        OOB estimate of  error rate: 1.05%
Confusion matrix:
     A    B    C    D    E class.error
A 2782    5    2    0    1 0.002867384
B   15 1869   15    0    0 0.015797788
C    0   15 1689    7    0 0.012857978
D    0    1   27 1579    1 0.018034826
E    0    0    5    9 1790 0.007760532
")
```
